Briefing paper on Ethnicity and COVID-19 status in Scotland

### 2020-07-01 Covid_ethnicity_Scotland_reportPartB.Rmd
Part B of ethnicity report - disaggregated ethnicity groupings.

@@

rownames(table.hospitalized.demog.ethnicsmrB) <-
  gsub("_", " ", rownames(table.hospitalized.demog.ethnicsmrB))
knitr::kable(table.hospitalized.demog.ethnicsmrB,
             escape=FALSE, 
             booktabs=TRUE,
             label="ethnicsmr.hosp",
             col.names=c(clean.header(colnames(table.hospitalized.demog.ethnicsmrB)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6), 
             caption="Associations of hospitalized disease with ethnicity in those with ethnicity recorded in Scottish Morbidity record") %>%
  kable_styling(latex_options="hold_position") %>% 
  column_spec(column=1, width="2.7cm") %>%
  column_spec(column=2, width="2cm")  %>%
  column_spec(column=3, width="1.85cm")  %>%
  column_spec(column=4, width="2.5cm") %>%
  column_spec(column=5, width="1.5cm") %>%
  column_spec(column=6, width="2.5cm") %>%
  column_spec(column=7, width="1.5cm")

@@

rownames(table.ethnic9smrB) <- gsub("_", " ", rownames(table.ethnic9smrB))
knitr::kable(table.ethnic9smrB,
             escape=FALSE, 
             booktabs=TRUE,
             label="ethnicsmr",
             col.names=c(clean.header(colnames(table.ethnic9smrB)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6), 
             caption="Associations of severe disease with ethnicity in those with ethnicity recorded in Scottish Morbidity record") %>%
  kable_styling(latex_options="hold_position") %>% 
  column_spec(column=1, width="2.7cm") %>%
  column_spec(column=2, width="1.85cm")  %>%
  column_spec(column=3, width="1.85cm")  %>%
  column_spec(column=4, width="2.5cm") %>%
  column_spec(column=5, width="1.5cm") %>%
  column_spec(column=6, width="2.8cm") %>%
  column_spec(column=7, width="1.5cm")
```

@@

rownames(table.hospitalized.demogB) <- gsub("_", " ", rownames(table.hospitalized.demogB))
knitr::kable(table.hospitalized.demogB,
             escape=FALSE, 
             booktabs=TRUE,
             label="onomap.hosp",
             col.names=c(clean.header(colnames(table.hospitalized.demogB)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6), 
             caption="Associations of hospitalized disease with demographic variables including ethnicity based on name classification") %>%
  add_header_above(c(" "=3, "Univariate"=2, "Multivariate"=2))  %>%
  kable_styling(latex_options="hold_position") %>% 
  column_spec(column=1, width="2.7cm") %>%
  column_spec(column=2, width="2cm")  %>%
  column_spec(column=3, width="1.85cm")  %>%
  column_spec(column=4, width="2.5cm") %>%
  column_spec(column=5, width="1.5cm") %>%
  column_spec(column=6, width="2.5cm") %>%
  column_spec(column=7, width="1.5cm")

@@

rownames(table.severe.demogB) <- gsub("_", " ", rownames(table.severe.demogB))
knitr::kable(table.severe.demogB,
             escape=FALSE, 
             booktabs=TRUE,
             label="onomap",
             col.names=c(clean.header(colnames(table.severe.demogB)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6),
             caption="Associations of severe disease with demographic variables including ethnicity based on name classification") %>%
  kable_styling(latex_options="hold_position") %>% 
  column_spec(column=1, width="2cm") %>%
  add_header_above(c(" "=3, "Univariate"=2, "Multivariate"=2)) %>%
  column_spec(column=2, width="1.85cm")  %>%
  column_spec(column=3, width="1.85cm")  %>%
  column_spec(column=4, width="2.5cm") %>%
  column_spec(column=5, width="1.5cm") %>%
  column_spec(column=6, width="2.5cm") %>%
  column_spec(column=7, width="1.5cm")

@@

rownames(table.severe.demogNRSB) <- gsub("_", " ", rownames(table.severe.demogNRSB))
knitr::kable(table.severe.demogNRSB,
             escape=FALSE, 
             booktabs=TRUE,
             label="onomapNRSB",
             col.names=c(clean.header(colnames(table.severe.demogNRSB)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6),
             caption="Associations of severe disease OR any COVID-19 death with demographic variables including ethnicity based on name classification") %>%
    kable_styling(latex_options="hold_position") %>% 
    column_spec(column=1, width="2cm") %>%
    add_header_above(c(" "=3, "Univariate"=2, "Multivariate"=2)) %>%
    column_spec(column=2, width="1.85cm")  %>%
    column_spec(column=3, width="1.85cm")  %>%
    column_spec(column=4, width="2.5cm") %>%
    column_spec(column=5, width="1.5cm") %>%
    column_spec(column=6, width="2.5cm") %>%
    column_spec(column=7, width="1.5cm")

>>>>

Where there is evidence of an elevated risk of "severe test positive disease" or "severe test positive or death" among those of South Asian ethnicity in Part A of this report, the observed rate ratios are slightly higher among those of Indian ethnicity compared to those of Pakistani/Bangladeshi using the disaggregated SMR approach.  For those with Black ethnicity in Part A of the report, the disaggregated data show that those classed as Black Caribbean have an increased risk of "severe test positive disease" and "severe test positive or death", as do Black Africans, but numbers are very small in these groups and therefore estimates are more unstable.

When applying the disaggregated ONOMAP approach, we see that those with Muslim South Asian ethnicity are at an increased risk of "severe test positive disease" or "severe test positive or death" and not those classed as Other South Asian ethnicity.

@@

rownames(table.hospitalized.demog.ethnicsmrB) <-
  gsub("_", " ", rownames(table.hospitalized.demog.ethnicsmrB))
knitr::kable(table.hospitalized.demog.ethnicsmrB,
             escape=FALSE, 
             booktabs=TRUE,
             label="ethnicsmr.hosp",
             col.names=c(clean.header(colnames(table.hospitalized.demog.ethnicsmrB)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6), 
             caption="Associations of hospitalized or severe disease among test positives with ethnicity in those with ethnicity recorded in Scottish Morbidity record") %>%
  kable_styling(latex_options="hold_position") %>% 
  column_spec(column=1, width="2.7cm") %>%
  column_spec(column=2, width="2cm")  %>%
  column_spec(column=3, width="1.85cm")  %>%
  column_spec(column=4, width="2.5cm") %>%
  column_spec(column=5, width="1.5cm") %>%
  column_spec(column=6, width="2.5cm") %>%
  column_spec(column=7, width="1.5cm")

@@

rownames(table.ethnic9smrB) <- gsub("_", " ", rownames(table.ethnic9smrB))
knitr::kable(table.ethnic9smrB,
             escape=FALSE, 
             booktabs=TRUE,
             label="ethnicsmr",
             col.names=c(clean.header(colnames(table.ethnic9smrB)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6), 
             caption="Associations of severe disease among test positives with ethnicity in those with ethnicity recorded in Scottish Morbidity record") %>%
  kable_styling(latex_options="hold_position") %>% 
  column_spec(column=1, width="2.7cm") %>%
  column_spec(column=2, width="1.85cm")  %>%
  column_spec(column=3, width="1.85cm")  %>%
  column_spec(column=4, width="2.5cm") %>%
  column_spec(column=5, width="1.5cm") %>%
  column_spec(column=6, width="2.8cm") %>%
  column_spec(column=7, width="1.5cm")
```

@@

rownames(table.hospitalized.demogB7) <- gsub("_", " ", rownames(table.hospitalized.demogB7))
knitr::kable(table.hospitalized.demogB7,
             escape=FALSE, 
             booktabs=TRUE,
             label="onomap.hosp",
             col.names=c(clean.header(colnames(table.hospitalized.demogB7)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6), 
             caption="Associations of hospitalized or severe disease among test positives with demographic variables including ethnicity based on name classification") %>%
  add_header_above(c(" "=3, "Univariate"=2, "Multivariate"=2))  %>%
  kable_styling(latex_options="hold_position") %>% 
  column_spec(column=1, width="2.7cm") %>%
  column_spec(column=2, width="2cm")  %>%
  column_spec(column=3, width="1.85cm")  %>%
  column_spec(column=4, width="2.5cm") %>%
  column_spec(column=5, width="1.5cm") %>%
  column_spec(column=6, width="2.5cm") %>%
  column_spec(column=7, width="1.5cm")

@@

#rownames(table.severe.demogB) <- gsub("_", " ", rownames(table.severe.demogB))
#knitr::kable(table.severe.demogB,
#             escape=FALSE, 
#             booktabs=TRUE,
#             label="onomapB",
#             col.names=c(clean.header(colnames(table.severe.demogB)[1:2]),
#                         "Rate ratio (95\\% CI)", "p-value",
#                         "Rate ratio (95\\% CI)", "p-value"),
#             align=rep("c", 6),
#             caption="Associations of severe disease with demographic variables including ethnicity based on name classification") %>%
#  kable_styling(latex_options="hold_position") %>% 
#  column_spec(column=1, width="2cm") %>%
#  add_header_above(c(" "=3, "Univariate"=2, "Multivariate"=2)) %>%
#  column_spec(column=2, width="1.85cm")  %>%
#  column_spec(column=3, width="1.85cm")  %>%
#  column_spec(column=4, width="2.5cm") %>%
#  column_spec(column=5, width="1.5cm") %>%
#  column_spec(column=6, width="2.5cm") %>%
#  column_spec(column=7, width="1.5cm")

```{r onomapB7, echo=FALSE, warning=FALSE, message=FALSE}

rownames(table.severe.demogB7) <- gsub("_", " ", rownames(table.severe.demogB7))
knitr::kable(table.severe.demogB7,
             escape=FALSE, 
             booktabs=TRUE,
             label="onomapB7",
             col.names=c(clean.header(colnames(table.severe.demogB7)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6),
             caption="Associations of severe disease among test positives with demographic variables including ethnicity based on name classification") %>%
  kable_styling(latex_options="hold_position") %>% 
  column_spec(column=1, width="2cm") %>%
  add_header_above(c(" "=3, "Univariate"=2, "Multivariate"=2)) %>%
  column_spec(column=2, width="1.85cm")  %>%
  column_spec(column=3, width="1.85cm")  %>%
  column_spec(column=4, width="2.5cm") %>%
  column_spec(column=5, width="1.5cm") %>%
  column_spec(column=6, width="2.5cm") %>%
  column_spec(column=7, width="1.5cm")

@@

#rownames(table.severe.demogNRSB) <- gsub("_", " ", rownames(table.severe.demogNRSB))
#knitr::kable(table.severe.demogNRSB,
#             escape=FALSE, 
#             booktabs=TRUE,
#             label="onomapNRSB",
#             col.names=c(clean.header(colnames(table.severe.demogNRSB)[1:2]),
#                         "Rate ratio (95\\% CI)", "p-value",
#                         "Rate ratio (95\\% CI)", "p-value"),
#             align=rep("c", 6),
#             caption="Associations of severe disease OR any COVID-19 death with demographic variables including ethnicity based on name #classification") %>%
#    kable_styling(latex_options="hold_position") %>% 
#    column_spec(column=1, width="2cm") %>%
#    add_header_above(c(" "=3, "Univariate"=2, "Multivariate"=2)) %>%
#    column_spec(column=2, width="1.85cm")  %>%
#    column_spec(column=3, width="1.85cm")  %>%
#    column_spec(column=4, width="2.5cm") %>%
#    column_spec(column=5, width="1.5cm") %>%
#    column_spec(column=6, width="2.5cm") %>%
#    column_spec(column=7, width="1.5cm")

```{r onomapNRSB7, echo=FALSE, warning=FALSE, message=FALSE}

rownames(table.severe.demogNRSB7) <- gsub("_", " ", rownames(table.severe.demogNRSB7))
knitr::kable(table.severe.demogNRSB7,
             escape=FALSE, 
             booktabs=TRUE,
             label="onomapNRSB7",
             col.names=c(clean.header(colnames(table.severe.demogNRSB7)[1:2]),
                         "Rate ratio (95\\% CI)", "p-value",
                         "Rate ratio (95\\% CI)", "p-value"),
             align=rep("c", 6),
             caption="Associations of severe disease OR any COVID-19 death with demographic variables including ethnicity based on name classification") %>%
    kable_styling(latex_options="hold_position") %>% 
    column_spec(column=1, width="2cm") %>%
    add_header_above(c(" "=3, "Univariate"=2, "Multivariate"=2)) %>%
    column_spec(column=2, width="1.85cm")  %>%
    column_spec(column=3, width="1.85cm")  %>%
    column_spec(column=4, width="2.5cm") %>%
    column_spec(column=5, width="1.5cm") %>%
    column_spec(column=6, width="2.5cm") %>%
    column_spec(column=7, width="1.5cm")
